---
title: "A Zoom Into Data-Related Jobs"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    favicon: favicon.png
    theme:
      version: 4
      bootswatch: minty
---


# Univariate Analysis {data-orientation=rows}

## Row {.tabset .tabset-fade}

### Numerical Variables {data-height=500}

```{r}
plot(1:10, main="CHART A")

plot(1:10, main="CHART B")

```

*** 

Some commentary about Frame 1.

### Categorical Variables {data-height=500}

```{r}
plot(1:10, main="CHART A")

plot(1:10, main="CHART B")


```

# Bivariate Analysis

## Row {.tabset .tabset-fade}

### Salary Vs. Company Location


```{r}
plot(1:10, main="CHART A")

```

### Salary Vs. Job Title

```{r}
plot(1:10, main="CHART A")

```

### Salary Vs. Experience Level

```{r}
plot(1:10, main="CHART A")

plot(1:10, main="CHART B")

```

# Further Analysis

## Row {.tabset .tabset-fade}

### Entry Level Salaries Vs. Job Titles \| IN EUROPE \|

```{r}
plot(1:10, main="CHART A")
```

### Entry Level Salaries Vs. Company Location \| IN EUROPE \|

```{r}
plot(1:10, main="CHART A")

plot(1:10, main="CHART B")
```
