---
title: "A Zoom Into Data-Related Jobs"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    favicon: favicon.png
    theme:
      version: 4
      bootswatch: minty
---
```{r The Whole Code, include=FALSE}
# Loading Everything I need

#Loading Packages
library(tidyverse)
library(ggplot2)
library(dplyr)
library(sf)
library(rnaturalearth)
library(scales)
library(viridis) 
library(wordcloud)
library(flexdashboard)

# Importing the dataset from a local path in my computer
data <- read.csv2("C:/Users/Samaher/Documents/DSE_R_Project/data.csv")

# Converting it into a tibble
data <- as.tibble(data)

# Replacing abbreviations with full descriptions in the 'experience_level' column
data$experience_level <- ifelse(data$experience_level == "EN", "entry level",
                           ifelse(data$experience_level == "EX", "executive",
                           ifelse(data$experience_level == "MI", "mid-level",
                           ifelse(data$experience_level == "SE", "senior", data$experience_level))))

# Replacing abbreviations with full descriptions in the 'employment_type' column
data$employment_type <- ifelse(data$employment_type == "CT", "contract",
                               ifelse(data$employment_type == "FL", "freelance",
                                      ifelse(data$employment_type == "FT", "full-time",
                                             ifelse(data$employment_type == "PT", "part-time", data$employment_type))))

# Given exchange rate
exchange_rate <- 0.92

# Converting USD salaries to EUR
data$salary_in_usd <- data$salary_in_usd * exchange_rate

# Changing column name to better reflect this change
names(data)[names(data) == "salary_in_usd"] <- "salary_in_eur"

# Removing the salary and salary_currency columns
data <- subset(data, select = -c(salary, salary_currency))

# Defining a mapping of country codes to country names for company locations
cl_country_mapping <- c(
  AE = "United Arab Emirates",
  AL = "Albania",
  AM = "Armenia",
  AR = "Argentina",
  AS = "American Samoa",
  AT = "Austria",
  AU = "Australia",
  BA = "Bosnia and Herzegovina",
  BE = "Belgium",
  BO = "Bolivia",
  BR = "Brazil",
  BS = "Bahamas",
  CA = "Canada",
  CF = "Central African Republic",
  CH = "Switzerland",
  CL = "Chile",
  CN = "China",
  CO = "Colombia",
  CR = "Costa Rica",
  CZ = "Czech Republic",
  DE = "Germany",
  DK = "Denmark",
  DZ = "Algeria",
  EE = "Estonia",
  EG = "Egypt",
  ES = "Spain",
  FI = "Finland",
  FR = "France",
  GB = "United Kingdom",
  GH = "Ghana",
  GR = "Greece",
  HK = "Hong Kong",
  HN = "Honduras",
  HR = "Croatia",
  HU = "Hungary",
  ID = "Indonesia",
  IE = "Ireland",
  IN = "India",
  IQ = "Iraq",
  IR = "Iran",
  IT = "Italy",
  JP = "Japan",
  KE = "Kenya",
  LT = "Lithuania",
  LU = "Luxembourg",
  LV = "Latvia",
  MA = "Morocco",
  MD = "Moldova",
  MK = "North Macedonia",
  MT = "Malta",
  MX = "Mexico",
  MY = "Malaysia",
  NG = "Nigeria",
  NL = "Netherlands",
  NZ = "New Zealand",
  PH = "Philippines",
  PK = "Pakistan",
  PL = "Poland",
  PR = "Puerto Rico",
  PT = "Portugal",
  RO = "Romania",
  RU = "Russia",
  SE = "Sweden",
  SG = "Singapore",
  SI = "Slovenia",
  SK = "Slovakia",
  TH = "Thailand",
  TR = "Turkey",
  UA = "Ukraine",
  US = "United States",
  VN = "Vietnam"
)

# Replacing values in the 'company_location' column using the mapping created above
data$company_location <- cl_country_mapping[data$company_location]

# Defining a mapping of country codes to country names for employee residence

er_country_mapping <- c(
  AE = "United Arab Emirates",
  AM = "Armenia",
  AR = "Argentina",
  AS = "American Samoa",
  AT = "Austria",
  AU = "Australia",
  BA = "Bosnia and Herzegovina",
  BE = "Belgium",
  BG = "Bulgaria",
  BO = "Bolivia",
  BR = "Brazil",
  CA = "Canada",
  CF = "Central African Republic",
  CH = "Switzerland",
  CL = "Chile",
  CN = "China",
  CO = "Colombia",
  CR = "Costa Rica",
  CY = "Cyprus",
  CZ = "Czech Republic",
  DE = "Germany",
  DK = "Denmark",
  DO = "Dominican Republic",
  DZ = "Algeria",
  EE = "Estonia",
  EG = "Egypt",
  ES = "Spain",
  FI = "Finland",
  FR = "France",
  GB = "United Kingdom",
  GH = "Ghana",
  GR = "Greece",
  HK = "Hong Kong",
  HN = "Honduras",
  HR = "Croatia",
  HU = "Hungary",
  ID = "Indonesia",
  IE = "Ireland",
  IN = "India",
  IQ = "Iraq",
  IR = "Iran",
  IT = "Italy",
  JE = "Jersey",
  JP = "Japan",
  KE = "Kenya",
  KW = "Kuwait",
  LT = "Lithuania",
  LU = "Luxembourg",
  LV = "Latvia",
  MA = "Morocco",
  MD = "Moldova",
  MK = "North Macedonia",
  MT = "Malta",
  MX = "Mexico",
  MY = "Malaysia",
  NG = "Nigeria",
  NL = "Netherlands",
  NZ = "New Zealand",
  PH = "Philippines",
  PK = "Pakistan",
  PL = "Poland",
  PR = "Puerto Rico",
  PT = "Portugal",
  RO = "Romania",
  RS = "Serbia",
  RU = "Russia",
  SE = "Sweden",
  SG = "Singapore",
  SI = "Slovenia",
  SK = "Slovakia",
  TH = "Thailand",
  TN = "Tunisia",
  TR = "Turkey",
  UA = "Ukraine",
  US = "United States",
  UZ = "Uzbekistan",
  VN = "Vietnam"
)

# Replacing values in the 'employee_residence' column using the mapping created above
data$employee_residence <- er_country_mapping[data$employee_residence]


salary_distribution <- ggplot(data, aes(x = salary_in_eur)) +
    geom_density(fill = "#5bc8af", color = "#202060") +
    labs(title = "Density Plot of Salary in EUR", x = "Salary in EUR", y = "Density") +
    scale_x_continuous(labels = scales::comma, breaks = seq(0, 500000, by = 100000)) # Modifying the x-axis scale to display breaks at intervals of 100,000 with comma formatting

remote_distribution <- ggplot(data, aes(x = remote_ratio)) +
    geom_density(fill = "#5bc8af", color = "#202060") +
    labs(title = "Density Plot of Remote Ratio", x = "Remote Ratio", y = "Density")

# Creating a table of company sizes and their frequency
company_sizes_freq <- table(data$company_size)

# Converting the table to a data frame
company_sizes_df <- data.frame(company_size = names(company_sizes_freq), frequency = as.numeric(company_sizes_freq))

# Defining a custom color palette (this step is optional, but, they're my favorite colors)
custom_palette <- c("#5bc8af","#202060", "#b030b0","#6c91bf", "#602080")

# Creating the pie chart for company sizes with percentages
company_sizes_pie <- ggplot(company_sizes_df, aes(x = "", y = frequency, fill = company_size)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar("y", start = 0) +
  labs(title = "Company Size Distribution", fill = "Company Size") +
  theme_void() +
  scale_fill_manual(values = custom_palette) +
  geom_text(aes(label = paste0(round((frequency/sum(frequency)) * 100), "%")),
            position = position_stack(vjust = 0.5), color = "white" )

# Creating a table of work years and their frequency
work_years_freq <- table(data$work_year)

# Converting the table to a data frame
work_years_df <- data.frame(work_year = names(work_years_freq), frequency = as.numeric(work_years_freq))

# Creating the pie chart for work years with percentages
work_year_pie <- ggplot(work_years_df, aes(x = "", y = frequency, fill = work_year)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar("y", start = 0) +
  labs(title = "Work Year Distribution", fill = "Work Year") +
  theme_void() +
  scale_fill_manual(values = custom_palette) +
  geom_text(aes(label = paste0(round((frequency/sum(frequency)) * 100), "%")),
            position = position_stack(vjust = 0.5), color = "white" ) 

# Creating a table of experience levels and their frequency
experience_levels_freq <- table(data$experience_level)

# Converting the table to a data frame
experience_levels_df <- data.frame(experience_level = names(experience_levels_freq), frequency = as.numeric(experience_levels_freq))

# Creating the pie chart for experience levels with percentages
exp_level_pie <- ggplot(experience_levels_df, aes(x = "", y = frequency, fill = experience_level)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar("y", start = 0) +
  labs(title = "Experience Level Distribution", fill = "Experience Level") +
  theme_void() +
  scale_fill_manual(values = custom_palette) +
  geom_text(aes(label = paste0(round((frequency/sum(frequency)) * 100), "%")),
            position = position_stack(vjust = 0.5), color = "white" )

# Creating a table of employment types and their frequency
employment_types_freq <- table(data$employment_type)

# Converting the table to a data frame
employment_types_df <- data.frame(employment_type = names(employment_types_freq), frequency = as.numeric(employment_types_freq))

# Creating the pie chart for employment types with percentages
emp_type_pie <- ggplot(employment_types_df, aes(x = "", y = frequency, fill = employment_type)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar("y", start = 0) +
  labs(title = "Employment Type Distribution", fill = "Employment Type") +
  theme_void() +
  scale_fill_manual(values = custom_palette) +
  geom_text(aes(label = paste0(round((frequency/sum(frequency)) * 100), "%")),
            position = position_stack(vjust = 0.5), color = "white" )

# Creating a df that has the company locations along with their freq
company_location_df <- data %>%
  group_by(company_location) %>%
  summarise(frequency = n())

# Getting world map data
world <- ne_countries(scale = "medium", returnclass = "sf")

# Merging company_location_df with world map data
map_data <- merge(world, company_location_df, by.x = "name", by.y = "company_location", all.x = TRUE)

# Plotting
company_location_map <- ggplot() +
  geom_sf(data = map_data, aes(fill = log(frequency + 1))) +
  scale_fill_gradient(low = "#5bc8af", high = "#202060") +
  labs(title="Company Location Distribution") +
  theme_minimal()

# Creating a data frame with employee residence locations and their frequency
employee_residence_df <- data %>%
  group_by(employee_residence) %>%
  summarise(frequency = n())

# Getting world map data
world <- ne_countries(scale = "medium", returnclass = "sf")

# Merging employee_residence_df with world map data
map_data_residence <- merge(world, employee_residence_df, by.x = "name", by.y = "employee_residence", all.x = TRUE)

# Plotting
employee_residence_map <- ggplot() +
  geom_sf(data = map_data_residence, aes(fill = log(frequency + 1))) +
  scale_fill_gradient(low = "#5bc8af", high = "#202060") +
  labs(title="Employee Residence Distribution") +
  theme_minimal()

# Creating a table of job titles and their frequency
job_titles_freq <- table(data$job_title)

# Converting the table to a data frame and sort by frequency
job_titles_df <- data.frame(job_title = names(job_titles_freq), frequency = as.numeric(job_titles_freq))
job_titles_df <- job_titles_df[order(job_titles_df$frequency, decreasing = TRUE), ]

# Calculating percentages
job_titles_df$percentage <- (job_titles_df$frequency / sum(job_titles_df$frequency)) * 100

# Selecting the top 10 most occurring job titles
top_10_jobs <- head(job_titles_df, 10)


# Creating a bar plot for the top 10 job titles with percentages on the y-axis
top_10_jobs_freq <- ggplot(top_10_jobs, aes(x = reorder(job_title, -frequency), y = percentage)) +
  geom_bar(stat = "identity", fill = "#5bc8af") +
  labs(title = "Top 10 Most Occurring Job Titles", x = "Job Titles", y = "Percentage") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Creating a dataframe with company locations and average salary for each country
salary_location_df <- data %>%
  group_by(company_location) %>%
  summarise(avg_salary = mean(salary_in_eur, na.rm = TRUE)) %>% # Calculating the average salary
  arrange(desc(avg_salary))  # Sorting by average salary in descending order

# Getting world map data
world <- ne_countries(scale = "medium", returnclass = "sf")

# Merging company_location_df with world map data
map_data <- merge(world, salary_location_df, by.x = "name", by.y = "company_location", all.x = TRUE)

# Plotting
salary_location_map <- ggplot() +
  geom_sf(data = map_data, aes(fill = avg_salary)) +
  scale_fill_gradient(low = "#5bc8af", high = "#202060") +
  labs(title="Average Salary Vs. Company Location") +
  theme_minimal()

# Selecting the top 10 countries
top_10_countries <- head(salary_location_df, 10)

# Creating a bar plot for top 10 countries with highest average salaries
top_10_countries_salary <- ggplot(top_10_countries, aes(x = reorder(company_location, -avg_salary), y = avg_salary)) +
  geom_bar(stat = "identity", fill = "#5bc8af") +
  labs(title = "Top 10 Countries with Highest Average Salaries",
       x = "Company Location", y = "Average Salary") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Creating a dataframe with job titles and their average salary
salary_jobtitle_df <- data %>%
  group_by(job_title) %>%
  summarise(avg_salary = mean(salary_in_eur, na.rm = TRUE))  

# Sorting by average salary in descending order to get the top salaries at the top
salary_jobtitle_df <- salary_jobtitle_df[order(-salary_jobtitle_df$avg_salary),]

# Selecting the top 10 job titles
top_10_jobtitles <- head(salary_jobtitle_df, 10)

# Creating a bar plot for the top 10 job titles
top_10_jobtitles_salary <- ggplot(top_10_jobtitles, aes(x = reorder(job_title, -avg_salary), y = avg_salary)) +
  geom_bar(stat = "identity", fill = "#5bc8af") +
  labs(title = "Top 10 Job Titles by Average Salary", x = "Job Title" , y = "Average Salary" ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::comma, breaks = pretty_breaks(n = 10))  # Formatting the Y-axis labels

# Reordering experience_level from entry level to executive
data$experience_level <- factor(data$experience_level, levels = c("entry level", "mid-level", "senior", "executive"))

# Creating boxplot for salaries across experience levels with custom colors
salary_exp_boxplot <- ggplot(data, aes(x = experience_level, y = salary_in_eur, fill = experience_level)) +
  geom_boxplot() +
  labs(title = "Salaries across Experience Levels", x = "Experience Level", y = "Salary in EUR") +
  scale_fill_manual(values = custom_palette) +  # Adding my favorite colors
  theme_minimal() +
  scale_y_continuous(labels = scales::comma, breaks = pretty_breaks(n = 10))

# Calculate sample size for each experience level
sample_size <- data %>% 
  group_by(experience_level) %>%     
  summarize(num = n())               

# Joining the datasets based on the experience_level column
data_with_size <- left_join(data, sample_size, by = "experience_level") %>%
  mutate(myaxis = paste0(experience_level, "\n", "n=", num)) # Creating a new column 'myaxis' combining experience level and sample size

# Plot
salary_exp_violin <- ggplot(data_with_size, aes(x = myaxis, y = salary_in_eur, fill = experience_level)) +
  geom_violin(width = 0.7, position = position_dodge(width = 0.6)) +
  geom_boxplot(width = 0.1, color = "white", alpha = 0.2, position = position_dodge(width = 0.6)) +
  scale_fill_manual(values = custom_palette) +  # Adding my favorite colors
  theme_minimal() +
  ggtitle("Salaries across Experience Levels") +
  xlab("") +
  ylab("Salary in EUR")+
  scale_y_continuous(labels = scales::comma, breaks = pretty_breaks(n = 10))

# Creating a vector of European countries
european_countries <- c("Spain", "France", "Germany", "Italy", "United Kingdom", "Netherlands", "Sweden", "Norway", "Finland", "Denmark", "Belgium", "Austria", "Greece", "Switzerland", "Portugal", "Ireland", "Czech Republic", "Romania", "Hungary", "Poland", "Slovakia", "Croatia", "Bulgaria", "Estonia", "Lithuania", "Latvia", "Slovenia", "Luxembourg", "Malta", "Cyprus")

# Filtering data for entry-level roles in Europe
entry_level_europe <- data[data$experience_level == "entry level" & data$company_location %in% european_countries, ]

# Calculating average salary for each job title in entry-level roles in Europe
avg_salary_entry_level_europe <- aggregate(salary_in_eur ~ job_title, entry_level_europe, FUN = mean)

# Sorting by average salary in descending order
sorted_data_europe <- avg_salary_entry_level_europe[order(-avg_salary_entry_level_europe$salary_in_eur), ]

# Selecting top 10 job titles for entry level in Europe
top_10_jobtitles_entry_europe <- head(sorted_data_europe, 10)

# Creating the bar plot
top_10_jobttitle_entry_eu <- ggplot(top_10_jobtitles_entry_europe, aes(x = reorder(job_title, salary_in_eur), y = salary_in_eur)) +
  geom_bar(stat = "identity", fill = "#5bc8af") +
  labs(title = "10 Highest-Paying Entry Level Data Roles in Europe", x = "Job Title", y = "Average Salary") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()  # To display the bars horizontally


# Calculating average salary for each country
avg_salary_entry_level_europe <- entry_level_europe %>%
  group_by(company_location) %>%
  summarise(avg_salary = mean(salary_in_eur, na.rm = TRUE)) %>%
  arrange(desc(avg_salary))  # Sorting by average salary in descending order

# Getting world map data for Europe
europe_map <- ne_countries(scale = "medium", continent = "Europe", returnclass = "sf")

# Merge salary data with the map data for Europe
map_data_europe <- merge(europe_map, avg_salary_entry_level_europe, by.x = "name", by.y = "company_location", all.x = TRUE)

# Plotting map of Europe showing average salaries for entry level
countries_entry_salary_eu <- ggplot() +
  geom_sf(data = map_data_europe, aes(fill = avg_salary)) +
  scale_fill_gradient(low = "#5bc8af", high = "#202060") +
  labs(title = "Average Salary for Entry Level in EU Countries") +
  theme_minimal()

# Sorting the data by average salary in descending order
top_10_countries_eu <- head(avg_salary_entry_level_europe[order(-avg_salary_entry_level_europe$avg_salary), ], 10)

# Creating a bar plot for the top 10 European countries with the highest average salaries
top_10_eu_salary_entry <- ggplot(top_10_countries_eu, aes(x = reorder(company_location, avg_salary), y = avg_salary)) +
  geom_bar(stat = "identity", fill = "#5bc8af") +
  labs(title = "Top 10 EU Countries with Highest Avg Salaries (Entry Level)", x = "Country", y = "Average Salary") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()

```

# Univariate Analysis

## Row {.tabset .tabset-fade}

### Salary Distribution

```{r}
plot(salary_distribution)
```

### Remote Ratio Distribution

```{r}
plot(remote_distribution)
```

### Company Sizes

```{r}
plot(company_sizes_pie)
```

### Work Years

```{r}
plot(work_year_pie)
```

### Experience Levels

```{r}
plot(exp_level_pie)
```

### Employment Types

```{r}
plot(emp_type_pie)
```

# Bivariate Analysis

## Row {.tabset .tabset-fade}

### Salary Vs. Company Location

```{r}
plot(salary_location_map)
```

### Salary Vs. Job Title

```{r}
plot(top_10_jobtitles_salary)
```

### Salary Vs. Experience Level

```{r}
plot(salary_exp_boxplot)

```

# Further Analysis

## Row {.tabset .tabset-fade}

### Entry Level Salaries Vs. Job Titles \| IN EUROPE \|

```{r}
plot(top_10_jobttitle_entry_eu)
```

### Entry Level Salaries Vs. Company Location \| IN EUROPE \|

```{r}
plot(top_10_eu_salary_entry)
```
