---
title: "A Zoom Into Data-Related Jobs"
output: html_notebook
---




**Name**: Samaher Brahem 

**Course**: Coding for DS & DM - R Module 

**Professor**: Giancarlo Manzi


### **1. Introduction**

This study aims to analyze salary differences among Data Scientists and other Data-related roles across different countries. Utilizing a dataset containing diverse employment types, experience levels, and company sizes, the study intends to conduct statistical analyses and visualizations using R. The objective is to identify potential salary disparities and trends across various global contexts. The study aims to shed light on nuanced salary patterns within the data science domain, offering insights into the variations of compensation among data-related roles on a global scale.

### **2. Data Used**

The *Salary of Data Scientists* dataset will be used for this analysis. It is made available on [Kaggle](https://www.kaggle.com/datasets/piyushborhade/salary-of-data-scientists) by the user, Piyush Borhade.

**A. Features of the dataset**:

- `work_year`: The year the salary was paid.

- `experience_level`: The level of work experience during the year

- `employment_type`: the type of employment for the role

- `job_title`: The job in which you worked during the year.

- `salary`: The total gross amount of salary paid.

- `salary_currency`: The currency of the salary paid as an ISO 4217 currency code.

- `salary_in_usd`: The salary in USD

- `employee_residence`: the country of primary residence of the employee during the employment year as an ISO 3166 country code.

- `remote_ratio`: The total amount of work done remotely

- `company_location`: The country of headquarters of the employer or contracting branch

- `company_size`: The average number of people who worked for the company during the year


**B. Limits of the dataset**:

1. No Shared Sources:

The information about how the data was collected and who collected it isn't clear. This lack of clarity raises doubts about how dependable the data is.


2. Geographical Bias:

The dataset primarily comprises data from the United States, resulting in an imbalance where other countries are underrepresented. This skew might limit the depth and accuracy of analyses, potentially leading to an overemphasis on American-centric insights and a lack of comprehensive understanding regarding other regions or global perspectives.


### **3. Data Processing**

#### **3. 1 Installing packages and opening libraries:**

- tidyverse

```{r}
library(tidyverse)

```

#### **3.2 Importing the dataset:**

```{r}
# Importing the dataset from a local path in my computer
data <- read.csv("C:/Users/Samaher/Documents/DSE_R_Project/data.csv")

# Converting it into a tibble
data <- as.tibble(data)

```


#### **3.3 Previewing the dataset:**

```{r}
# Understanding the structure of the dataset
str(data)
```

```{r}
# Displaying a sample of the data
head(data)
```


#### **3.4 Cleaning and formatting the dataset:**

##### **3.4.1 Experience Level:**

There are 4 experience levels in this dataset:

- EN = Entry Level

- MI = Mid-Level

- SE = Senior

- EX = Executive


```{r}
# Checking how many experience levels are included in this dataset
table(data$experience_level)
```
Let's make those levels understandable by anyone who reads them!

```{r}
# Replacing abbreviations with full descriptions in the 'experience_level' column
data$experience_level <- ifelse(data$experience_level == "EN", "entry level",
                           ifelse(data$experience_level == "EX", "executive",
                           ifelse(data$experience_level == "MI", "mid-level",
                           ifelse(data$experience_level == "SE", "senior", data$experience_level))))

# Checking the updated 'experience_level' column
table(data$experience_level)
```

##### **3.4.2 Employment Type:**

There are 4 employment types in this dataset:

- FT = full-time

- PT = part-time

- CT = contract

- FL = freelance

```{r}
table(data$employment_type)
```
I will replace those abbreviations with the full word for better readability.

```{r}
# Replacing abbreviations with full descriptions in the 'employment_type' column
data$employment_type <- ifelse(data$employment_type == "CT", "contract",
                               ifelse(data$employment_type == "FL", "freelance",
                                      ifelse(data$employment_type == "FT", "full-time",
                                             ifelse(data$employment_type == "PT", "part-time", data$employment_type))))

# Checking the updated 'employment_type' column
table(data$employment_type)
```

##### **3.4.3 Salaries:**

When it comes to salaries, we can see that there are 3 columns: 
1. `salary`
2. `salary_currency`
3. `salary_in_usd` 

The salaries, stored in `salary` column, are collected in local currencies, mentioned in `salary_currency` column. Then, all those salaries were converted to USD and stored in `salary_in_usd`. 

```{r}
table(data$salary_currency)
```

For the purpose of this study, it would be better to have the salaries in EUR. That's why I will replace the `salary_in_usd` column with `salary_in_eur` and convert the salaries to EUR. I will also remove the `salary` and `salary_currecny` columns since they are no longer needed. 

The exchange rate for the conversion from USD to EUR is as follows:

1 United States Dollar equals
0.92 Euro

P.S. The date of this conversion is: Dec 16, 09:12 UTC


```{r}
# Given exchange rate
exchange_rate <- 0.92

# Converting USD salaries to EUR
data$salary_in_usd <- data$salary_in_usd * exchange_rate

# Changing column name to better reflect this change
names(data)[names(data) == "salary_in_usd"] <- "salary_in_eur"
```

```{r}
# Removing the salary and salary_currency columns
data <- subset(data, select = -c(salary, salary_currency))
```

##### **3.4.4 Countries:**

###### **3.4.4.1 Company Location:**

The column `company_location` reflects the countries where the companies are established. Those countries are written as abbreviations. While some country abbreviations are widely known, there are others that are less known. 

```{r}
table(data$company_location)
```


To ensure clarity, I decided to replace the abbreviations with the full country names.

```{r}
# Defining a mapping of country codes to country names for company locations
cl_country_mapping <- c(
  AE = "United Arab Emirates",
  AL = "Albania",
  AM = "Armenia",
  AR = "Argentina",
  AS = "American Samoa",
  AT = "Austria",
  AU = "Australia",
  BA = "Bosnia and Herzegovina",
  BE = "Belgium",
  BO = "Bolivia",
  BR = "Brazil",
  BS = "Bahamas",
  CA = "Canada",
  CF = "Central African Republic",
  CH = "Switzerland",
  CL = "Chile",
  CN = "China",
  CO = "Colombia",
  CR = "Costa Rica",
  CZ = "Czech Republic",
  DE = "Germany",
  DK = "Denmark",
  DZ = "Algeria",
  EE = "Estonia",
  EG = "Egypt",
  ES = "Spain",
  FI = "Finland",
  FR = "France",
  GB = "United Kingdom",
  GH = "Ghana",
  GR = "Greece",
  HK = "Hong Kong",
  HN = "Honduras",
  HR = "Croatia",
  HU = "Hungary",
  ID = "Indonesia",
  IE = "Ireland",
  IN = "India",
  IQ = "Iraq",
  IR = "Iran",
  IT = "Italy",
  JP = "Japan",
  KE = "Kenya",
  LT = "Lithuania",
  LU = "Luxembourg",
  LV = "Latvia",
  MA = "Morocco",
  MD = "Moldova",
  MK = "North Macedonia",
  MT = "Malta",
  MX = "Mexico",
  MY = "Malaysia",
  NG = "Nigeria",
  NL = "Netherlands",
  NZ = "New Zealand",
  PH = "Philippines",
  PK = "Pakistan",
  PL = "Poland",
  PR = "Puerto Rico",
  PT = "Portugal",
  RO = "Romania",
  RU = "Russia",
  SE = "Sweden",
  SG = "Singapore",
  SI = "Slovenia",
  SK = "Slovakia",
  TH = "Thailand",
  TR = "Turkey",
  UA = "Ukraine",
  US = "United States",
  VN = "Vietnam"
)

# Replacing values in the 'company_location' column using the mapping created above
data$company_location <- cl_country_mapping[data$company_location]

```


###### **3.4.4.2 Employee Residence:**

Now, we'll do the same process for `employee_residence` column. 

```{r}
#Checking which countries are included
table(data$employee_residence)
```
```{r}
# Defining a mapping of country codes to country names for employee residence

er_country_mapping <- c(
  AE = "United Arab Emirates",
  AM = "Armenia",
  AR = "Argentina",
  AS = "American Samoa",
  AT = "Austria",
  AU = "Australia",
  BA = "Bosnia and Herzegovina",
  BE = "Belgium",
  BG = "Bulgaria",
  BO = "Bolivia",
  BR = "Brazil",
  CA = "Canada",
  CF = "Central African Republic",
  CH = "Switzerland",
  CL = "Chile",
  CN = "China",
  CO = "Colombia",
  CR = "Costa Rica",
  CY = "Cyprus",
  CZ = "Czech Republic",
  DE = "Germany",
  DK = "Denmark",
  DO = "Dominican Republic",
  DZ = "Algeria",
  EE = "Estonia",
  EG = "Egypt",
  ES = "Spain",
  FI = "Finland",
  FR = "France",
  GB = "United Kingdom",
  GH = "Ghana",
  GR = "Greece",
  HK = "Hong Kong",
  HN = "Honduras",
  HR = "Croatia",
  HU = "Hungary",
  ID = "Indonesia",
  IE = "Ireland",
  IN = "India",
  IQ = "Iraq",
  IR = "Iran",
  IT = "Italy",
  JE = "Jersey",
  JP = "Japan",
  KE = "Kenya",
  KW = "Kuwait",
  LT = "Lithuania",
  LU = "Luxembourg",
  LV = "Latvia",
  MA = "Morocco",
  MD = "Moldova",
  MK = "North Macedonia",
  MT = "Malta",
  MX = "Mexico",
  MY = "Malaysia",
  NG = "Nigeria",
  NL = "Netherlands",
  NZ = "New Zealand",
  PH = "Philippines",
  PK = "Pakistan",
  PL = "Poland",
  PR = "Puerto Rico",
  PT = "Portugal",
  RO = "Romania",
  RS = "Serbia",
  RU = "Russia",
  SE = "Sweden",
  SG = "Singapore",
  SI = "Slovenia",
  SK = "Slovakia",
  TH = "Thailand",
  TN = "Tunisia",
  TR = "Turkey",
  UA = "Ukraine",
  US = "United States",
  UZ = "Uzbekistan",
  VN = "Vietnam"
)

# Replacing values in the 'employee_residence' column using the mapping created above
data$employee_residence <- er_country_mapping[data$employee_residence]

```



### **4. Data Analysis**

### **5. Conclusion**

